Amplitude metrics for cellular \\ circadian bioluminescence reporters

### 2014-09-02 response_letter.tex

The authors start by describing phase response curves PRCs, ARCs and phase density equations.
They also describe practical approaches to computing these.
Using such a computed perturbed distribution, the authors show how to calculate an ``average oscillation'' of the population.

My main concern is that I cannot tell what is novel in the manuscript.
Since references are scant in the main body of the text, this is somewhat obscure.
For example, phase response curves have been discussed in numerous places, as have amplitude response curves.
The phase density methods and an associated Fokker-Planck equation (Eq.~31) also date back at least to Kuramoto and Stratonovich, with a lot of work following that.
For instance, the ``synchronization index'', $\rho$, is the order parameter in Kuramoto's treatment, and has been studied in hundreds of papers.
I guess this is well known, but some references should be given.
Stochastic oscillators at the population level, with a focus on PRCs, have been studied in a series of papers by Tanaka, Goldobin, Ermentrout and others - see also the standard, but somewhat dated book by Pikovsky, Rosenblum and Kurths.

Pages 4-14 of the manuscript therefore contain a lot of known results.
Some of the results may be new - for instance, I have not seen Eq.~(26) before - but the majority of the results presented here have been standard fare in the voluminous literature on stochastic oscillators.
Is the novelty here that these can be used to analyze data? If so, this is not well explained.
\end{reviewer}

We acknowledge that many of the mathematical formulas presented in the ``Results'' section have been used by many other studies: our intention was to introduce concepts and notations to readers who might be less familiar with the large body of previous literature.
The reviewer's point is a good one, however, and we have endeavored to draw as clear a line as possible between new concepts and derivations and those which come from the literature.
{\itshape\bfseries To this end, we have moved as much of the introductory derivations as possible to the ``Materials and Methods'' section}, introducing previous work on phase-diffusion models before proceeding with sensitivity-based methods for finding the ARC.
Additionally, we have included additional references for historical derivations of the results:
\begin{manuscript}{Pages 5-6}
  The synchrony of the population can be modeled using a probability density function $p(\theta, \tilde{t})$ that describes the probability of finding an oscillator at each phase {\bfseries (33)}.
The usefulness of probability density functions in describing the phase and amplitude responses of populations of circadian cells has been {\bfseries previously shown} (14).
As with all probability density functions:
\begin{equation}
  \int_0^{2\pi} p(\theta, \tilde{t}) \; d\theta = 1
  \tag{12}
\end{equation}
The shape of $p(\theta, \tilde{t})$ changes as the cells advance in time.
Stochastic effects cause the population to gradually desynchronize as slight cycle-to-cycle deviations are propagated throughout the population {\bfseries (34). For a infinite population of oscillators,} these effects are well-described by a {\bfseries Fokker-Plank} equation (35):
\end{manuscript}

\begin{manuscript}{Page 21}
\bfseries
33. Kuramoto, Y., 1984. Chemical Oscillations, Waves, and Turbulence. Springer-Verlag Berlin
Heidelberg.\\
34. Teramae, J.-n., and D.~Tanaka, 2004. {Robustness of the noise-induced phase synchronization in a general
  class of limit cycle oscillators.} \emph{Phys. Rev. Lett.} 93:204103.
\end{manuscript}

### 2014-09-04 main_text.tex
Adding ackowledgement to anonymous reviewers

\section*{Acknowledgments}
We thank Yongqiang Wang for the helpful discussions.
This work was supported by the National Institutes of Health/National Institute of General Medical Sciences under award number 1R01GM096873-01 and by the Institute for Collaborative Biotechnologies through grant W911NF-09-0001 from the U.S.
Army Research Office.

>>>>

\section*{Acknowledgments}
We thank Yongqiang Wang for the helpful discussions, and gratefully acknowledge the three anonymous reviewers whose comments/suggestions helped improve this work. 
This work was supported by the National Institutes of Health/National Institute of General Medical Sciences under award number 1R01GM096873-01 and by the Institute for Collaborative Biotechnologies through grant W911NF-09-0001 from the U.S.\ Army Research Office.
